Results

### 2025-08-13 notebooks/000_data.qmd
update

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SPEDAS: tplot, set_meta
using GLMakie
```

@@

@@

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        [Wind.V_GSE_3DP, Wind.V_GSE_K0],
        [Wind.n_p_K0, Wind.n_p_3DP, Wind.n_p_nonlin],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

>>>>

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SPEDAS: tplot, set_meta
using Speasy
using GLMakie
```

@@

Different coordinate frames

```{julia}
tmin = DateTime("2021-01-15")
tvars = [
    "PSP_SWP_SPI_SF00_L3_MOM/VEL_RTN_SUN",
    "PSP_SWP_SPI_SF00_L3_MOM/VEL_SC",
    "PSP_SWP_SPI_SF00_L3_MOM/VEL_INST",
    "PSP_SWP_SPI_SF00_L3_MOM/MAGF_SC",
    "PSP_FLD_L2_MAG_SC_4_SA_PER_CYC/psp_fld_l2_mag_SC_4_Sa_per_Cyc"
]
tplot(tvars, tmin, tmin + Hour(1))
```

"PSP_SWP_SPI_SF00_L3_MOM" ~3.5s resolution

@@

```{julia}
let tmin = DateTime("2021-01-15"), add_unit = false
    tvars = (
        Wind.V_GSE_K0,
        Wind.V_GSE_3DP,
        [
            SpeasyProduct("WI_K0_SWE/Np"; labels=["Proton"]),
            SpeasyProduct("WI_PM_3DP/P_DENS"; labels=["Proton (3DP, PESA Low)"]),
            SpeasyProduct("WI_H1_SWE/Proton_Np_nonlin", labels=["Proton (non-linear fitting)"]),
            SpeasyProduct("WI_H1_SWE/Alpha_Na_nonlin", labels=["Alpha (non-linear fitting)"]),
            Wind.n_e
        ],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

### 2025-08-13 notebooks/11_spectra.qmd
update

```{julia}
using Dates
using PSPAnisotropy: Wind
tmin = DateTime("2007-07-06")
tmax = DateTime("2007-07-08")
B = Wind.B(tmin, tmax; add_unit=false)
```

```{julia}
using DSP

pxx = periodogram(B[:, 1])

@@

```{julia}
# first, add the package if you haven't already:
# import Pkg; Pkg.add("Multitaper")

using Statistics
using LinearAlgebra
using Multitaper

const μ0 = 4π * 1e-7     # vacuum permeability [H/m]

"""
    compute_fluctuations(v, B, rho)

Given
  • v    : N×3 array of velocity (m/s)
  • B    : N×3 array of magnetic field (T)
  • rho  : length‐N vector of mass density (kg/m³)

Returns
  • δv   : N×3 array of velocity fluctuations
  • δb   : N×3 array of magnetic‐field fluctuations in Alfvén units (m/s)
"""
function compute_fluctuations(v::AbstractMatrix, B::AbstractMatrix, rho::AbstractVector)
    @assert size(v) == size(B) "v and B must be the same size"
    @assert size(v, 1) == length(rho) "rho length must match rows of v,B"

    # subtract mean
    v̄ = mean(v, dims=1)
    δv = v .- v̄

    # magnetic field in velocity units: b = B / sqrt(μ0 ρ)
    b = B ./ sqrt.(μ0 .* rho)[:, None]
    b̄ = mean(b, dims=1)
    δb = b .- b̄

    return δv, δb
end

"""
    trace_psd_mt(x, dt; NW=4)

Compute the trace (sum over 3 components) PSD of a 3‐component fluctuation array x via multitaper.

Arguments:
  • x    : N×3 array of fluctuation time series
  • dt   : sampling interval [s]
Keyword:
  • NW   : time–bandwidth product (default 4)

Returns
  • f    : vector of frequency bins [Hz]
  • P    : vector of trace power spectrum P(f) = ∑ᵢ S_ii(f)
"""
function trace_psd_mt(x::AbstractMatrix, dt::Real; NW::Real=4)
    N, nc = size(x)
    @assert nc == 3 "x must be N×3"

    fs = 1 / dt
    # how many tapers? standard choice K = 2NW − 1
    K = Int(2 * NW - 1)

    # accumulate over components
    f = nothing
    P_trace = nothing

    for comp in 1:3
        # multitaper spectrum for this component
        spec = mt_spectrum(x[:, comp];
            NW=NW,
            K=K,
            fs=fs,
            adaptive=false)  # or true for adaptive weighting

        f === nothing && (f = spec.f)
        P_trace === nothing && (P_trace = zeros(length(f)))
        P_trace .+= spec.amp .^ 2           # mt_spectrum.amp is the amplitude spectrum
    end

    return f, P_trace
end

"""
    compute_sigma(v, B, rho, dt; NW=4)

>>>>

```{julia}
using Dates
using DSP
using PSPAnisotropy
using PSPAnisotropy: Wind
using SPEDAS
using Unitful

tmin = DateTime("2007-07-06")
tmax = DateTime("2007-07-08")
B = Wind.B_GSE(tmin, tmax)
V = Wind.V_GSE_3DP(tmin, tmax) |> dropna
n = Wind.n_p_3DP(tmin, tmax) |> dropna
```

```{julia}
using PlasmaFormulary

V_B = Alfven_velocity.(tsync(B, n)...)

z_pm = PlasmaFormulary.Elsässer.(tsync(V, B, n)...)
z_plus, z_minus = getindex.(z_pm, 1), getindex.(z_pm, 2)
```

```{julia}
"""
    trace_pgram(f, x; kwargs...)

Compute the trace (sum over components) power spectral density of a multi-component time series. Returns a `Periodogram` object with trace power spectral density
"""
function trace_pgram(f, x; fs=nothing, kwargs...)
    # Compute periodogram for each component and sum
    fs = something(fs, samplingrate(x) / 1u"Hz")
    pgrams = map(eachcol(x)) do col
        f(col; fs, kwargs...)
    end
    trace_power = sum(power, pgrams)
    # Return periodogram with trace power and same frequency grid
    return DSP.Periodograms.Periodogram(trace_power, freq(pgrams[1]))
end

@@

```{julia}
V_B_pxx_mt = trace_pgram(mt_pgram, ustrip(V_B))

V_pxx = trace_pgram(periodogram, ustrip(V .|> u"m/s"))
V_pxx_mt = trace_pgram(mt_pgram, ustrip(V .|> u"m/s"))
z_plus_pxx_mt = trace_pgram(mt_pgram, ustrip(z_plus))
z_minus_pxx_mt = trace_pgram(mt_pgram, ustrip(z_minus))
```

```{julia}
using CairoMakie

function plot_pxx_pxx_mt!(ax, pxx, pxx_mt; step=10, linewidth=2, alpha=0.8)
    # Get frequency and power data
    freqs_pxx = freq(pxx)
    powers_pxx = power(pxx)
    freqs_mt = freq(pxx_mt)
    powers_mt = power(pxx_mt)

    p1 = lines!(ax, freqs_pxx[2:step:end], powers_pxx[2:step:end]; label="Periodogram", linewidth, alpha)
    p2 = lines!(ax, freqs_mt[2:step:end], powers_mt[2:step:end]; label="Multitaper", linewidth, alpha)
    axislegend(ax, position=:lt)
    return p1, p2
end

function plot_pxx_pxx_mt(pxx, pxx_mt; kw...)
    f = Figure(size=(800, 600))
    ax = Axis(f[1, 1],
        xlabel="Frequency (Hz)",
        ylabel="Power Spectral Density",
        xscale=log10,
        yscale=log10,
        title="Comparison of Power Spectral Density Methods"
    )
    plot_pxx_pxx_mt!(ax, pxx, pxx_mt; kw...)
    return f
end

plot_pxx_pxx_mt(V_pxx, V_pxx_mt)
```

```{julia}
"""
    parker_spiral_direction(r_au; Ω=2.87e-6, V_sw=400e3)

Compute the expected Parker spiral magnetic field direction at distance r_au from the Sun.

Arguments:
- `r_au`: distance from Sun in AU
- `Ω`: solar rotation rate [rad/s] (default: 2.87e-6 rad/s)
- `V_sw`: solar wind speed [m/s] (default: 400 km/s)

Returns:
- Unit vector in Parker spiral direction (radial, tangential, normal components)
"""
function parker_spiral_direction(r_au; Ω=2.87e-6, V_sw=400e3)
    r = r_au * 1.496e11  # convert AU to meters

    # Parker spiral angle
    φ = atan(Ω * r / V_sw)

    # Unit vector: (radial, tangential, normal)
    B_parker = [cos(φ), -sin(φ), 0.0]
    return B_parker / norm(B_parker)
end

"""
    rectify_magnetic_field(B, r_au=1.0)

Rectify magnetic field direction based on Parker spiral to ensure consistent
z+ (outward) and z- (inward) Alfvén wave identification.

Arguments:
- `B`: N×3 magnetic field array
- `r_au`: distance from Sun in AU (default: 1.0)

Returns:
- `B_rect`: rectified magnetic field with consistent Parker spiral alignment
"""
function rectify_magnetic_field(B, r_au=1.0)
    B_parker = parker_spiral_direction(r_au)
    B_rect = copy(B)

    for i in 1:size(B, 1)
        B_vec = B[i, :]
        # Check if B is aligned with Parker spiral (dot product > 0)
        if dot(B_vec, B_parker) < 0
            # Flip B direction if anti-aligned with Parker spiral
            B_rect[i, :] = -B_vec
        end
    end

    return B_rect
end
```

```{julia}
# Plot comparison of Elsasser variable spectra
f = Figure(size=(1200, 800))

step = 10

ax3 = Axis(f[1, 1],
    xlabel="Frequency (Hz)",
    ylabel="Power Spectral Density",
    xscale=log10,
    yscale=log10
)

ax4 = Axis(f[1, 2],
    xlabel="Frequency (Hz)",
    ylabel="Power Spectral Density",
    xscale=log10,
    yscale=log10
)

plot_spectras!(ax3, V_B_pxx_mt, V_pxx_mt; labels=["B", "V"], step=step, linewidth=2, alpha=0.8)
axislegend(ax3, position=:lt)
xlims!(ax3, (1e-4, 5e-1))
ylims!(ax3, (1e6, 1e13))

plot_spectras!(ax4, z_plus_pxx_mt, z_minus_pxx_mt; labels=["z+ (Outward)", "z- (Inward)"], step=step, linewidth=2, alpha=0.8)
axislegend(ax4, position=:lt)
xlims!(ax4, (1e-4, 5e-1))
ylims!(ax4, (1e6, 1e13))

f
```

The spectra of residual energy Er, cross helicity Ec, total energy Et, normalized residual energy σr, normalized cross helicity σc, Alfv ́en ratio rA, and Elsasser ratio rE were then calculated,

$$
\begin{gathered}
E_{\mathrm{r}}=E_{\mathrm{v}}-E_{\mathrm{b}}, \\
E_{\mathrm{c}}=E_{+}-E_{-}, \\
E_{\mathrm{t}}=E_{\mathrm{v}}+E_{\mathrm{b}}, \\
\sigma_{\mathrm{r}}=\frac{E_{\mathrm{v}}-E_{\mathrm{b}}}{E_{\mathrm{v}}+E_{\mathrm{b}}}=\frac{r_{\mathrm{A}}-1}{r_{\mathrm{A}}+1}, \\
\sigma_{\mathrm{c}}=\frac{E_{+}-E_{-}}{E_{+}+E_{-}}=\frac{r_{\mathrm{E}}-1}{r_{\mathrm{E}}+1}, \\
r_{\mathrm{A}}=\frac{E_{\mathrm{v}}}{E_{\mathrm{b}}}=\frac{1+\sigma_{\mathrm{r}}}{1-\sigma_{\mathrm{r}}}, \\
r_{\mathrm{E}}=\frac{E_{+}}{E_{-}}=\frac{1+\sigma_{\mathrm{c}}}{1-\sigma_{\mathrm{c}}} .
\end{gathered}
$$

```{julia}
"""
    compute_sigma(v, B, rho, dt; NW=4)
